---
title: "BC Seagrass Metacommunity Analysis- new Hmsc-R package"
author: "Keila Stark"
date: "20/06/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Install and load Packages

```{r packages, echo = FALSE, warning=FALSE, message=FALSE}

install.packages('devtools')
install.packages('Rcpp')
install.packages('RcppArmadillo')
install.packages('coda')
install.packages('beanplot')
install.packages('circlize')
install.packages('corrplot')
install.packages('coda')
install.packages('dendextend')

library(devtools)
library(HMSC)
library(vegan)
library(tidyverse)
library(viridis)
library(corrplot)
library(RColorBrewer)
library(dendextend)

install_github("hmsc-r/HMSC", build_opts = c("--no-resave-data", "--no-manual"))
library(Hmsc)

```


# Read data

```{r read data, echo=FALSE, warning=FALSE, message=FALSE}
setwd("~/Github/HMS-Seagrass/Data")

Y <- read.csv("Y_matrix.csv", header = TRUE, stringsAsFactors = FALSE)
Y <- Y[-c(103,104),]
rownames(Y) <- Y[,1]

X <- read.csv("X_matrix.csv", header = TRUE, stringsAsFactors = FALSE) # 
rownames(X) <- X[,1]

spatial <- read.csv("spatial_matrix.csv", header = TRUE, stringsAsFactors = FALSE) # lat&long coordinates
rownames(spatial) <- spatial[,1]


studyDesign <- read.csv("pi.csv", header = TRUE, stringsAsFactors = FALSE) #factors

Y <- Y[,-1] #remove actual site names from Y matrix now that rownames for all datasets have been established
X <- X[,-1]


## Remove NA's / get into right format
X[is.na(X)] <- 0
Y[is.na(Y)] <- 0
Y <- as.matrix(Y)
X <- as.data.frame(X)
spatial <- as.data.frame(spatial)
spatial$longitude <- spatial$longitude-min(spatial$longitude)

spatial <- data.frame(spatial = sprintf('spatial_%.2d',1:102)) #spatial factor column for studyDesign
studyDesign <- cbind(studyDesign, spatial)
studyDesign$spat <- factor(studyDesign$Quadrat)

rL1 = HmscRandomLevel(units = studyDesign$Quadrat)
rL2 = HmscRandomLevel(units = studyDesign$Site)
rL3 = HmscRandomLevel(units = studyDesign$Region)
rL4= HmscRandomLevel(sData = spatial)
```




# Construct HMSC model object and fit with MCMC

```{r fit model, echo=FALSE, warning=FALSE, message=FALSE}


hM <- Hmsc(Y = Y, XData = X, XFormula = ~eelgrass_lai + temprange + tempmean + nitratemean + curveloc + salinitymean + dissoxmean + macroalgae_biomass + detritus_biomass + shoot_density, studyDesign = studyDesign, ranLevels = list(Quadrat = rL1, Site = rL2, Region = rL3, spat = rL4), distr = "poisson")
 
# Run MCMC chains. 

mod_V2 <- sampleMcmc(hM, samples = 200000 , transient = 1000, thin = 100, verbose = 10)

```

# Evaluate MCMC convergence
We ensured MCMC chains were well-mixed for parameters beta (species niches/ responses to environmental covariates) and omega (residual species-to-species covariances). While beyond the scope of this study, parameters for phylogenetic signal and traits are also enabled in the HMSC framework.

```{r evaluate markov chains}
# computationally intense so #'d out these lines
mpost <- convertToCodaObject(mod)
par(mfrow=c(3,2))
ess.beta = effectiveSize(mpost$Beta)
psrf.beta = gelman.diag(mpost$Beta,multivariate=FALSE)$psrf
hist(ess.beta)
hist(psrf.beta)
#ess.gamma = effectiveSize(mpost$Gamma)
#psrf.gamma = gelman.diag(mpost$Gamma,multivariate=FALSE)$psrf 
hist(ess.gamma)
#hist(psrf.gamma)
#sppairs = matrix(sample(x = 1:ns^2, size = 100))
#tmp = mpost$Omega[[1]]
#for (chain in 1:length(tmp)){
#   tmp[[chain]] = tmp[[chain]][,sppairs]
#}
#ess.omega = effectiveSize(tmp)
#psrf.omega = gelman.diag(tmp,multivariate=FALSE)$psrf hist(ess.omega)
#hist(psrf.omega)
```


```{r}
par(mar = c(8,14,1,4))
postBeta = getPostEstimate(mod, parName = "Beta")
plotBeta(mod, post = postBeta, param = "Support", supportLevel = 0.95)
```

```{r variation partitioning}
partition <- createPartition(mod, nfolds = 2)
partition <- createPartition(mod1, nfolds = 2)
preds_fullmodel <- computePredictedValues(mod, partition = partition)
preds_no_env <- computePredictedValues(mod1, partition = partition)

MF <- evaluateModelFit(hM = mod, predY = preds)

#variance partitioning
VP <- computeVariancePartitioning(mod)
VP <- computeVariancePartitioning(mod, group = c(1,2,3,3,4,5,6,7,8,9,10), groupnames = c("Eelgrass LAI" ,"Temp range" ,"Mean temperature", "Mean nitrates" ,"Max current velocity" , "Mean salinity" , "Mean dissolved O2", "Macroalgae biomass", "Detritus biomass", "Eelgrass shoot density")) 

VP.test <- computeVariancePartitioning(mod)

plotVariancePartitioning(mod, VP = VP)# use default function from hmsc package to see mean variance explained for each covariate, then print manually below...

VP.df <- as.data.frame(VP$vals) %>% 
  mutate(effect = factor(c("Eelgrass LAI (0.7%)", "Temp range (17%)","Mean temperature (9.8%)",  "Mean nitrates (8.7%)","Max current velocity (9%)","Mean salinity (4.8%)", "Mean dissolved O2 (12.9%)", "Macroalgae biomass (1%)","Detritus biomass (0.8%)","Eelgrass shoot density (2%)", "Random: Quadrat (10.7%)","Random: Site (9.5%)", "Random: Region (5.2%)","Random: Spatial distance (7.9%)"), levels = rev(c("Eelgrass LAI (0.7%)", "Temp range (17%)","Mean temperature (9.8%)",  "Mean nitrates (8.7%)","Max current velocity (9%)","Mean salinity (4.8%)", "Mean dissolved O2 (12.9%)", "Macroalgae biomass (1%)","Detritus biomass (0.8%)","Eelgrass shoot density (2%)", "Random: Quadrat (10.7%)","Random: Site (9.5%)", "Random: Region (5.2%)","Random: Spatial distance (7.9%)")), ordered = TRUE)) 

target <- c("Temp range (17%)","Mean temperature (9.8%)",  "Mean nitrates (8.7%)","Max current velocity (9%)","Mean salinity (4.8%)", "Mean dissolved O2 (12.9%)", "Macroalgae biomass (1%)","Detritus biomass (0.8%)","Eelgrass LAI (0.7%)","Eelgrass shoot density (2%)", "Random: Quadrat (10.7%)","Random: Site (9.5%)", "Random: Region (5.2%)","Random: Spatial distance (7.9%)") # re-ordering factors so all biotic/biometrics are together for ease of looking

VP.df$effect <- VP.df[match(target, VP.df$effect),]

VP.df$effect <- c("Temp range (17%)","Mean temperature (9.8%)",  "Mean nitrates (8.7%)","Max current velocity (9%)","Mean salinity (4.8%)", "Mean dissolved O2 (12.9%)", "Macroalgae biomass (1%)","Detritus biomass (0.8%)","Eelgrass LAI (0.7%)","Eelgrass shoot density (2%)", "Random: Quadrat (10.7%)","Random: Site (9.5%)", "Random: Region (5.2%)","Random: Spatial distance (7.9%)") # re-ordering factors so all biotic/biometrics are together for ease of looking

VP.df1 <- read.csv("vpdf.csv", header=TRUE)
colnames(VP.df1) <- spnames

VP.df1 <- VP.df1 %>%
  gather(key = taxon, value = Variance, -effect) 


VP.df1 <- VP.df1%>% 
  group_by(taxon) %>% 
  mutate(mean = Variance)

VP.df1$taxon <- factor(VP.df1$taxon, 
                        levels = colnames(mod$Y)[order(colSums(mod$Y),decreasing = TRUE)], ordered = TRUE)

R2.df <- data.frame(R2 = round(MF$SR2,1), taxon = colnames(mod$Y))

par(mar=c(8,8,2.5,13))
ggplot(VP.df1, aes(y = Variance, x = taxon, fill = factor(effect, levels = c("Temp range (17%)","Mean temperature (9.8%)",  "Mean nitrates (8.7%)","Max current velocity (9%)","Mean salinity (4.8%)", "Mean dissolved O2 (12.9%)", "Macroalgae biomass (1%)","Detritus biomass (0.8%)","Eelgrass LAI (0.7%)","Eelgrass shoot density (2%)", "Random: Quadrat (10.7%)","Random: Site (9.5%)", "Random: Region (5.2%)","Random: Spatial distance (7.9%)"))))+
  geom_bar(stat = "identity", color = 1)+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 75, hjust = 1))+
  scale_fill_manual(values = c("#F0F921FF","#F0F921FF","#F0F921FF","#F0F921FF","#F0F921FF","#F0F921FF", "#BB3754FF","#BB3754FF","#BB3754FF","#BB3754FF", "grey30" ,"grey40" ,"grey50" , "grey60"), name = "Mean variance explained across all species")+
  geom_text(data = R2.df, aes(y = -0.02, fill = NULL, label = R2), size = 2)+
  geom_point(data = R2.df, aes(y = -0.06, fill = NULL, size = R2))+
  scale_size_continuous(breaks = seq(0.15,0.60,by = 0.15))+
  xlab(label = "Species")

`````



# Species associations/ correlation plots
```{r}

OmegaCor = computeAssociations(mod) #compute omega parameter
supportLevel = 0.95

toPlot1 = ((OmegaCor[[1]]$support>supportLevel) # first-pass of quadrat level associations
+ (OmegaCor[[1]]$support<(1-supportLevel))>0)*OmegaCor[[1]]$mean

toPlot2 = ((OmegaCor[[2]]$support>supportLevel)
+ (OmegaCor[[2]]$support<(1-supportLevel))>0)*OmegaCor[[2]]$mean # first-pass of site-level associations

toPlot3 = ((OmegaCor[[3]]$support>supportLevel) #first-pass of region-level associations
+ (OmegaCor[[3]]$support<(1-supportLevel))>0)*OmegaCor[[3]]$mean
library(corrplot)

corrplot(toPlot, method = "color",col=colorRampPalette(c("#21908C","white","#440154"))(200),  tl.cex=.75, tl.col="black", title=paste("random effect level:", mod$rLNames[1]), mar=c(0,0,0,0))

corrplot(toPlot2, method = "color",col=colorRampPalette(c("#21908C","white","#440154"))(200),  tl.cex=.75, tl.col="black", title=paste("random effect level:", mod$rLNames[2]), mar=c(0,0,1,0))

corrplot(toPlot3, method = "color",col=colorRampPalette(c("#21908C","white","#440154"))(200), tl.cex=.75, tl.col="black", title=paste("random effect level:", mod$rLNames[3]), mar=c(0,0,1,0))

corrplot(toPlot, method = "color",col=colorRampPalette(c("#21908C","white","#440154"))(200), order = "hclust", hclust.method= "ward.D2", tl.cex=.75, tl.col="black", title=paste("random effect level:", mod$rLNames[1]), mar=c(0,0,1,0))

corrplot(toPlot2, method = "color",col=colorRampPalette(c("#21908C","white","#440154"))(200), order = "hclust" , hclust.method= "ward.D2", tl.cex=.75, tl.col="black", title=paste("random effect level:", mod$rLNames[2]), mar=c(0,0,1,0))

corrplot(toPlot3, method = "color",col=colorRampPalette(c("#21908C","white","#440154"))(200), order = "hclust" , hclust.method = "ward.D2", tl.cex=.75, tl.col="black", title=paste("random effect level:", mod$rLNames[3]), mar=c(0,0,1,0))
```

```{r}
# create custom functions for heatmaps
imagine1 <- function (comm, col = col, fill = TRUE, 
                      xlab = "", ylab = "", yline = 2, xline = 2, sitenames = rownames(comm), 
                      speciesnames = colnames(comm), binary = TRUE) 
{
  reverse <- nrow(comm):1
  comm <- comm[reverse, ]
  par(mar = c(1, 9, 11.2, 15))
  
  image(1:dim(comm)[2], 1:dim(comm)[1], t(comm), col = c("#00000000","#52525203","#52525205","#52525208","#5252520A","#5252520D","#5252520F","#52525212","#52525214","#52525217","#5252521A","#5252521C","#5252521F","#52525221","#52525224","#52525226","#52525229","#5252522B","#5252522E","#52525230","#52525233","#52525236","#52525238","#5252523B","#5252523D","#52525240","#52525242","#52525245","#52525247","#5252524A","#5252524D","#5252524F","#52525252","#52525254","#52525257","#52525259","#5252525C","#5252525E","#52525261","#52525263","#52525266","#52525269","#5252526B","#5252526E","#52525270","#52525273","#52525275","#52525278","#5252527A","#5252527D","#52525280","#52525282","#52525285","#52525287","#5252528A","#5252528C","#5252528F","#52525291","#52525294","#52525296","#52525299","#5252529C","#5252529E","#525252A1","#525252A3","#525252A6","#525252A8","#525252AD","#525252B0","#525252B3","#525252B5","#525252B8","#525252BA","#525252BD","#525252BF","#525252C2","#525252C4","#525252C7","#525252C9","#525252CC","#525252CF","#525252D1","#525252D4","#525252D6","#525252D9","#525252DB","#525252DE","#525252E0","#525252E3","#525252E6","#525252E8","#525252EB","#525252ED","#525252F0","#525252F2","#525252F5","#525252F7","#525252FA","#525252FC"), xlab = "", ylab = "", axes = FALSE)
  box()
  if (length(sitenames) > 1) {
    axis(2, at = 1:dim(comm)[1], labels = sitenames, las = 1, 
         cex.axis = 1, lwd.ticks = 0)
  }
  if (length(speciesnames) > 1) {
    axis(3, at = 1:dim(comm)[2], labels = speciesnames, las = 2, 
         cex.axis = 1, lwd.ticks = 0)
  }
  mtext(xlab, 3, cex = 1.5, line = xline)
  mtext(ylab, 2, cex = 1.5, line = yline)
}

imagine2 <- function (comm, col = col, fill = TRUE, 
                      xlab = "", ylab = "", yline = 2, xline = 2, sitenames = rownames(comm), 
                      speciesnames = colnames(comm), binary = TRUE, add = TRUE) 
{
  reverse <- nrow(comm):1
  comm <- comm[reverse, ]
  par(mar = c(1, 9, 11.2, 15))
  
  image(1:dim(comm)[2], 1:dim(comm)[1], t(comm), col = c("#00000000","#21908C03","#21908C05","#21908C08","#21908C0A","#21908C0D","#21908C0F","#21908C12","#21908C14","#21908C17","#21908C1A","#21908C1C","#21908C1F","#21908C21","#21908C24","#21908C26","#21908C29","#21908C2B","#21908C2E","#21908C30","#21908C33","#21908C36","#21908C38","#21908C3B","#21908C3D","#21908C40","#21908C42","#21908C45","#21908C47","#21908C4A","#21908C4D","#21908C4F","#21908C52","#21908C54","#21908C57","#21908C59","#21908C5C","#21908C5E","#21908C61","#21908C63","#21908C66","#21908C69","#21908C6B","#21908C6E","#21908C70","#21908C73","#21908C75","#21908C78","#21908C7A","#21908C7D","#21908C80","#21908C82","#21908C85","#21908C87","#21908C8A","#21908C8C","#21908C8F","#21908C91","#21908C94","#21908C96","#21908C99","#21908C9C","#21908C9E","#21908CA1","#21908CA3","#21908CA6","#21908CA8","#21908CAD","#21908CB0","#21908CB3","#21908CB5","#21908CB8","#21908CBA","#21908CBD","#21908CBF","#21908CC2","#21908CC4","#21908CC7","#21908CC9","#21908CCC","#21908CCF","#21908CD1","#21908CD4","#21908CD6","#21908CD9","#21908CDB","#21908CDE","#21908CE0","#21908CE3","#21908CE6","#21908CE8","#21908CEB","#21908CED","#21908CF0","#21908CF2","#21908CF5","#21908CF7","#21908CFA","#21908CFC"), xlab = "", ylab = "", axes = FALSE, add = TRUE)
  box()
  if (length(sitenames) > 1) {
    axis(2, at = 1:dim(comm)[1], labels = sitenames, las = 1, 
         cex.axis = 1, lwd.ticks = 0)
  }
  if (length(speciesnames) > 1) {
    axis(3, at = 1:dim(comm)[2], labels = speciesnames, las = 2, 
         cex.axis = 1, lwd.ticks = 0)
  }
  mtext(xlab, 3, cex = 1.5, line = xline)
  mtext(ylab, 2, cex = 1.5, line = yline)
}

imagine3 <- function (comm, col = col, fill = TRUE, 
                      xlab = "", ylab = "", yline = 2, xline = 2, sitenames = rownames(comm), 
                      speciesnames = colnames(comm), binary = TRUE, add = TRUE) 
{
  reverse <- nrow(comm):1
  comm <- comm[reverse, ]
  par(mar = c(1, 9, 11.2, 15))
  
  image(1:dim(comm)[2], 1:dim(comm)[1], t(comm), col = c("#00000000","#44015403","#44015405","#44015408","#4401540A","#4401540D","#4401540F","#44015412","#44015414","#44015417","#4401541A","#4401541C","#4401541F","#44015421","#44015424","#44015426","#44015429","#4401542B","#4401542E","#44015430","#44015433","#44015436","#44015438","#4401543B","#4401543D","#44015440","#44015442","#44015445","#44015447","#4401544A","#4401544D","#4401544F","#44015452","#44015454","#44015457","#44015459","#4401545C","#4401545E","#44015461","#44015463","#44015466","#44015469","#4401546B","#4401546E","#44015470","#44015473","#44015475","#44015478","#4401547A","#4401547D","#44015480","44015482","#44015487","#4401548A","#4401548C","#4401548F","#44015491","#44015494","#44015496","#44015499","#4401549C","#4401549E","#440154A1","#440154A3","#440154A6","#440154A8","#440154AD","#440154B0","#440154B3","#440154B5","#440154B8","#440154BA","#440154BD","#440154BF","#440154C2","#440154C4","#440154C7","#440154C9","#440154CC","#440154CF","#440154D1","#440154D4","#440154D6","#440154D9","#440154DB","#440154DE","#440154E0","#440154E3","#440154E6","#440154E8","#440154EB","#440154ED","#440154F0","#440154F2","#440154F5","#440154F7","#440154FA","#440154FC"), xlab = "", ylab = "", axes = FALSE, add = TRUE)
  box()
  if (length(sitenames) > 1) {
    axis(2, at = 1:dim(comm)[1], labels = sitenames, las = 1, 
         cex.axis = 1, lwd.ticks = 0)
  }
  if (length(speciesnames) > 1) {
    axis(3, at = 1:dim(comm)[2], labels = speciesnames, las = 2, 
         cex.axis = 1, lwd.ticks = 0)
  }
  mtext(xlab, 3, cex = 1.5, line = xline)
  mtext(ylab, 2, cex = 1.5, line = yline)
}

```


```{r}
# setup data for corrplot
yo <- cbind(studyDesign, Y)
yo1 <- yo %>%
  select(-Spatial, -Quadrat,-Region,-spat,-spatial)
yo2 <- yo1 %>%
  group_by(Site) %>%
  summarise_each(funs(sum))
  
yo3 <- yo2 %>%
  select(c(Spirorbis.sp,Leptochelia.sp,Harpacticoid1,Lacuna.spp,Nereis.sp,Aoroides.spp,Porcellidium.sp,Alvania.compacta,Caprella.laeviuscula,Unknown.Ischyoceridae.gammarid))

topten <- cbind(yo2$Site, yo3)
rownames(topten) <- topten$`yo2$Site`

topten <- topten %>%
  select(-`yo2$Site`)


#### extract 20 most abundant from toplots
quad_cor <- toPlot1[c(11,26,28,5,12,16,27,2,23,21,22,9,17,19,20,30,24,6,4,3),c(11,26,28,5,12,16,27,2,23,21,22,9,17,19,20,30,24,6,4,3)]
site_cor <- toPlot2[c(11,26,28,5,12,16,27,2,23,21,22,9,17,19,20,30,24,6,4,3),c(11,26,28,5,12,16,27,2,23,21,22,9,17,19,20,30,24,6,4,3)]
reg_cor <- toPlot3[c(11,26,28,5,12,16,27,2,23,21,22,9,17,19,20,30,24,6,4,3),c(11,26,28,5,12,16,27,2,23,21,22,9,17,19,20,30,24,6,4,3)]


par(mfrow=c(1,1))

corrplot(quad_cor, method = "color",col=colorRampPalette(c("#21908C","white","#440154"))(200), order = "hclust", hclust.method= "ward.D2", tl.cex=.75, tl.col="black")

corrplot(site_cor, method = "color",col=colorRampPalette(c("#21908C","white","#440154"))(200), order = "hclust" , hclust.method= "ward.D2", tl.cex=.75, tl.col="black")

corrplot(reg_cor, method = "color",col=colorRampPalette(c("#21908C","white","#440154"))(200), order = "hclust" , hclust.method = "ward.D2", tl.cex=.75, tl.col="black")


##hrowing the original code here in case i need it again..
corrplot(quad_cor, method = "color",col=colorRampPalette(c("#21908C","white","#440154"))(200), order = "hclust", hclust.method= "ward.D2", tl.cex=.3, tl.col="black", title=paste("random effect level:", mod$rLNames[1]), mar=c(0,0,1,0))
corrplot(site_cor, method = "color",col=colorRampPalette(c("#21908C","white","#440154"))(200), order = "hclust" , hclust.method= "ward.D2", tl.cex=.3, tl.col="black", title=paste("random effect level:", mod$rLNames[2]), mar=c(0,0,1,0))
corrplot(reg_cor, method = "color",col=colorRampPalette(c("#21908C","white","#440154"))(200), order = "hclust" , hclust.method = "ward.D2", tl.cex=.3, tl.col="black", title=paste("random effect level:", mod$rLNames[3]), mar=c(0,0,1,0))

#making new matrices for heatmap

blues <- matrix(nrow = nrow(topten), ncol=ncol(topten))
blues[,c(3,7,10)] <- topten[,c(3,7,10)]
blues[,-c(3,7,10)] <- 0
rownames(blues) <- rownames(topten)
colnames(blues) <- colnames(topten)

purples <- matrix(nrow = nrow(topten), ncol=ncol(topten))
purples[,c(1,2,5,9)] <- topten[,c(1,2,5,9)]
purples[,-c(1,2,5,9)] <- 0
rownames(purples) <- rownames(topten)
colnames(purples) <- colnames(topten)

greys <- matrix(nrow = nrow(topten), ncol=ncol(topten))
greys[,c(4,6,8)] <- topten[,c(4,6,8)]
greys[,-c(4,6,8)] <- 0
rownames(greys) <- rownames(topten)
colnames(greys) <- colnames(topten)
##### 
blues <- matrix(nrow = nrow(predo2), ncol=ncol(predo2))
blues[,c(3,7,10)] <- predo2[,c(3,7,10)]
blues[,-c(3,7,10)] <- 0
rownames(blues) <- rownames(predo2)
colnames(blues) <- colnames(predo2)

purples <- matrix(nrow = nrow(predo2), ncol=ncol(predo2))
purples[,c(1,2,5,9)] <-predo2[,c(1,2,5,9)]
purples[,-c(1,2,5,9)] <- 0
rownames(purples) <- rownames(predo2)
colnames(purples) <- colnames(predo2)

greys <- matrix(nrow = nrow(predo2), ncol=ncol(predo2))
greys[,c(4,6,8)] <- predo2[,c(4,6,8)]
greys[,-c(4,6,8)] <- 0
rownames(greys) <- rownames(predo2)
colnames(greys) <- colnames(predo2)


# Make cluster dendrogram with predicted site-level abundances 
pred.site <-read.csv("predsite.csv", header = TRUE, stringsAsFactors = FALSE)
predo <- cbind(studyDesign, preds)
predo <- as.matrix(predo)
predo1 <- predo %>%
  select(-Spatial, -Quadrat,-Region,-spat,-spatial)

predo3 <-predo1 %>%
  group_by(Site) %>%
  summarise_each(funs(sum)) %>%
  select(Spirorbis.sp,Leptochelia.sp,Harpacticoid1,Lacuna.spp,Nereis.sp,Aoroides.spp,Porcellidium.sp,Alvania.compacta,Caprella.laeviuscula,Unknown.Ischyoceridae.gammarid)

predo3 <- as.matrix(predo3)
rownames(predo3) <- rownames(topten)

par(mar = c(0.55, 26, 10.85, 0.35))
dendo  <- decostand(predo3, method = "normalize") %>% dist %>% hclust %>% as.dendrogram %>%   set("branches_k_color", k = 3, value = c("#440154","#440154","#21908C" )) %>% set("labels", NULL) %>%    set("by_labels_branches_lwd", value = c(1:10)) %>% plot_horiz.dendrogram(axes = FALSE) 

#now that we've seen how the dendrogram falls out, time to make the imagine feed-in matrix rows match that order so the sites correspond correctly
#target <- c("Sarita", "Robbers","Sidney_Spit","Lyall_Harbour","Elbow_Bank", "Ducking", "Triquet_North" , "Triquet_Bay" ,"Ramsay","Choked_Inner","Gallagher_Bay","Cabbage_Isl", "James_Bay","Sandspit", "Dodger_Channel","Indian", "Louscoone")

#predo2 <- read.csv("predicteds.csv")

par(new = TRUE) # this line is important or else heatmap will replace dendrogram

imagine1(greys) #old colour scheme, modelled sp counts
imagine2(blues)
imagine3(purples)
```

# FIG 1b : Distance-decay of community similarity with model, raw data, and model but with environmental covariates set to mean values
```{r} 
#distance-decay
library(vegan)
library(reshape)

mod1 <- mod 
mod1$X <- matrix(rep(colMeans(mod1$X),each=n),
                      nrow=n,ncol=ncol(mod1$X)) #create version of the model where env covariates set to mean value

partition <- createPartition(mod, nfolds = 4)
preds_fullmodel <- computePredictedValues(mod, partition = partition) 
partition <- createPartition(mod1, nfolds = 4)
preds_no_env <- computePredictedValues(mod1, partition = partition) #model with mean env covariates

rawdist <- vegdist(Y, method = "bray") # calculate BC dissimilarity for raw data
rawdistlong <- subset(melt(as.matrix(rawdist)), value!=0)

preds_fullmodel1 <- preds_fullmodel[,,mean(1:200)] #mean of 200 slices
predmat <- vegdist(preds_fullmodel1, method = "bray" ) #calculate BC dissimilarity for PREDICTED data
predmatlong <- subset(melt(as.matrix(predmat)), value!=0)

preds_no_env1 <- preds_no_env[,,mean(1:200)] #take mean of all slices
no_env_mat<- vegdist(preds_no_env1, method = "bray") #calculate BC dissimilarity for predicted data with no effect of covariates (spatial distance and random effects only)
no_env_mat_long <- subset(melt(as.matrix(no_env_mat)), value!=0)

#PREPARED GEOGRAPHIC/ SPATIAL DISTANCE MATRIX
spatdistmat <- vegdist(spatial, method = "euclidean")  #melt the spatial distance matrix
spatdistmat1 <- spatdistmat*111 #convert lat/long degrees to kilometres
spatdistlong1 <- subset(melt(as.matrix(spatdistmat1)), value!=0)
#corrdist <- distancedecay2$`Spatial distance`*111

#COMBINE ALL LONG FORM DISTANCE DATA INTO ONE DATA FRAME
distancedecay <- cbind(predmatlong, no_env_mat_long, rawdistlong, spatdistlong1)
distancedecay1 <- distancedecay[,c(3,6,7,8,9,12)] #remove redundant columns 
colnames(distancedecay1) <- c("hmsc_dissim", "hmsc_noenv_dissim", "Site1", "Site2","raw_dissim","Spatial distance") #rename columns for ease of understanding


fullpredlm <- lm(distancedecay1$hmsc_dissim~distancedecay1$`Spatial distance`)
noenvlm <- lm(distancedecay1$hmsc_noenv_dissim~distancedecay1$`Spatial distance`)
rawlm <- lm(distancedecay1$raw_dissim~distancedecay1$`Spatial distance`)
#abline(fullpredlm)
#anova(distdeclm)
#anova(avglm)
#abline(avglm)

col1 <- rgb(253,231,37, max = 255, alpha = 20) 
col2 <- rgb(33,144,140, max = 255, alpha = 20)
col3 <- rgb(68,1,84, max = 255, alpha = 20)

par(mar = c(7,6,1,1)) #setup graphics parameters 
par(oma = c(4, 1, 1, 1))
plot.new()

plot(distancedecay1$hmsc_dissim~jitter(distancedecay1$`Spatial distance`), pch = 1,cex = 2, col=col1, xlab = "", ylab = "")
mtext(side=1, line=2, "Pairwise distance between samples (km)", font=2,cex=1.2)
mtext(side=2, line=3, "Pairwise Bray-Curtis dissimilarity index",  font=2, cex=1.2)
points(jitter(distancedecay1$raw_dissim)~jitter(distancedecay1$`Spatial distance`),pch=3,col=col3,xaxt='n',yaxt="n")
points(jitter(distancedecay1$hmsc_noenv_dissim)~jitter(distancedecay1$`Spatial distance`),pch=2,col=col2,xaxt='n',yaxt="n")
abline(fullpredlm,col="#440154FF",lwd = 4, lty =2)
abline(noenvlm,col="#21908CFF",  lwd = 4, lty = 2)
abline(rawlm, col = "#FDE725FF", lwd = 4, lty = 1)
par(xpd = FALSE)
legend("bottomleft",inset = c(0.3,0.001), legend=c("Full model (Space + environmental covariates)","Spatial distance + random effects only (no effect of environment)", "Real data"),col=c("#440154FF","#21908CFF", "#FDE725FF") ,cex= 0.9, lty=c(2,2,1), lwd = 4, bty = "n")

distancedecay2 <- distancedecay1 %>%
  select("Spatial distance", "hmsc_dissim", "hmsc_noenv_dissim", "raw_dissim") %>%
  gather(-`Spatial distance`, key = "variable", value = "value")

ggplot(distancedecay2, aes(x = `Spatial distance`, y = value))+  geom_point(aes(shape = variable, color = variable), alpha = 0.05) + xlab("Pairwise spatial distance (km)") + geom_smooth(aes(color = variable, fill = variable), method = "lm", se = TRUE) +ylab("Bray-Curtis pairwise dissimilarity between samples")  + scale_color_viridis(discrete = TRUE, option = "D", labels = c("Full model (Space + environmental covariates)","Spatial autocorrelation/random effects only (no effect of environment)", "Real data")) + scale_fill_viridis(discrete = TRUE) + theme_classic() + theme(legend.title = element_blank()) + theme(legend.position = "none")

```


# FIGURE 1a : MAP OF STUDY SITES
```{r}
library(ggplot2)
library(maps)
library(mapdata)
library(grid)

wholespatial <- read.csv("wholesitespatial.csv", header = TRUE, stringsAsFactors = FALSE)
plot.new()
par(mar = c(0,0,0,0))
big <- map("worldHires", "Canada", xlim = c(-131.7,-123), ylim=c(48.9, 53) , fill = TRUE, col = "grey80")

par(new = TRUE)

rect(-125.5763-0.1, 49.11286-0.1, -126.1426+0.1, 49.19492+0.1, col=NULL, border="firebrick", xpd=FALSE) # clayoquot sound

rect(-125.1469-0.1, 48.83528-0.1, -125.08+0.1, 48.89597+0.1, col=NULL, border="firebrick", xpd=FALSE) # barkley sound


rect(-131.4826-0.1, 52.23015-0.1,  -131.2062+0.1, 52.56887+0.1, col=NULL, border="firebrick", xpd=FALSE) # haida gwaii

rect(-122.9-0.1, 48.44074-0.1, -123.6972+0.1, 48.94043+0.1, col=NULL, border="firebrick", xpd=FALSE) # gulf islands


rect(-128.12+0.1, 51.67-0.1, -128.25-0.2, 51.81+0.1, col=NULL, border="firebrick", xpd=FALSE) # calvert


points(spatial$longitude, spatial$latitude, pch=21, col="black", bg = "blue", cex=0.8) 

HG <-locator(1) #specify where i want labels and legend to be with my cursor
CS <-locator(1)
BS <-locator(1)
GI <-locator(1)
HK <- locator(1)
text(HG,"Haida Gwaii", cex = 0.75)
text(CS,"Clayoquot Sound", cex = 0.75)
text(BS,"Barkley Sound", cex = 0.75)
text(GI,"Gulf Islands", cex = 0.75)
text(HK,"Central Coast", cex = 0.75)

map.scale(x = -131.6, y = 48.5, ratio = FALSE, cex = 0.72)
```


#Gathering Bio-ORACLE covariates
```{r}
install.packages("sdmpredictors")
install.packages("leaflet")
# Load package
library(sdmpredictors)
library(leaflet)

# Explore datasets in the package
list_datasets()

# Explore layers in a dataset
list_layers("Bio-ORACLE")

# Download specific layers to the current directory
layers <- load_layers(layercodes = c("BO2_nitraterange_bdmax", "BO2_tempmean_bdmax", "BO2_temprange_bdmax", "BO2_curvelmax_bdmax", "BO2_nitratemean_bdmax", "BO2_salinityrange_bdmax", "BO2_salinitymean_bdmax", "BO_ph", "BO2_dissoxmean_bdmax", "BO2_dissoxrange_bdmax"),equalarea=FALSE, rasterstack=TRUE)

bathy <- load_layers("BO_bathymean")

# Visualise sites of interest in google maps
m <- leaflet()
m <- addTiles(m)
m <- addMarkers(m, lng=spatial$longitude, lat=spatial$latitude, popup=spatial$quad)
m

# Extract environmental values from layers
environment <- data.frame(Name=spatial$quad,depth=extract(bathy,spatial[,2:3]) , env = raster::extract(layers,spatial[,2:3]))
write.csv(environment, file = "env.csv")

# Check layer statistics
layer_stats()

# Check Pearson correlation coefficient between layers
layers_correlation()

```


